---
title: "My Fancy Presentation"
author: "Naqeeb Asif"
date: "7 March 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## My new slide
Some content

***
A slide without a title

## My title

### A top level heading

#### Second heading

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

## My new slide

Plain text

*italicized text*

**Bold text**

`for(i %in% 1:10)`

## My new slide

```{r echo=TRUE}
head(mtcars)
```

## My new slide
```{r comment="",fig.align='center',fig.cap='My plot'}
library(ggplot2)
ggplot(data=mtcars,aes(x=wt,y=mpg))+xlab("Weight") + ylab("MPG") +
  geom_point()
```