---
title: "Google Vis"
author: "Naqeeb"
date: "6 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Google Vis tutorial

## Line chart
library(googleVis)
op <- options(gvis.plot.tag = "chart")
## Add the mean
CityPopularity$Mean=mean(CityPopularity$Popularity)
CC <- gvisComboChart(CityPopularity, xvar='City',
          yvar=c('Mean', 'Popularity'),
          options=list(seriesType='bars',
                       width=450, height=300,
                       title='City Popularity',
                       series='{0: {type:"line"}}'))
plot(CC)
```

## Line chart with two axis

```{r}
suppressPackageStartupMessages(library(googleVis))
op <- options(gvis.plot.tag='chart')

df=data.frame(country=c("US", "GB", "BR"), 
              val1=c(10,13,14), 
              val2=c(23,12,32))

Line2 <- gvisLineChart(df, "country", c("val1","val2"),
                       options=list(
                         series="[{targetAxisIndex: 0},
                                 {targetAxisIndex:1}]",
                         vAxes="[{title:'val1'}, {title:'val2'}]"
                       ))

Line2
```

## Bar Chart
```{r}
suppressPackageStartupMessages(library(googleVis))
op <- options(gvis.plot.tag='chart')

df=data.frame(country=c("US", "GB", "BR"), 
              val1=c(10,13,14), 
              val2=c(23,12,32))
Bar <- gvisBarChart(df)
print(Bar,"chart")
```

```{r message=FALSE}
op <- options(gvis.plot.tag='chart')
Motion=gvisMotionChart(Fruits, 
                       idvar="Fruit", 
                       timevar="Year")
print(Motion,"chart")
```

