---
title: "Data Science Specialization Capstone Project - Week 2"
author: "Naqeeb Asif"
date: "4 December 2018"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Synapsis
This document describes the data acquisition, basic summaries, cleaning, 
pre-processing  and basic exploratory analysis done in correspondence to the 
requirements in Data Science Capstone Projects.

## Acquisition of data and basic Summary
The data is downloaded from the [link](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip)
provided in the instructions of the project. Following three US files are loaded
from the data provided:

* en_US.blogs.txt
* en_US.twitter.txt
* en_US.news.txt
```{r loadLibraries,echo=FALSE,warning=FALSE,message=FALSE}
library(tidyverse)
library(tidytext)
library(parallel)
library(pbapply)
# Calculate the number of cores
no_cores <- detectCores() - 1

# Initiate cluster
cl <- makeCluster(no_cores)
```

```{r acquisition,echo=FALSE,results=FALSE,cache=TRUE,message=FALSE,warning=FALSE}
ifelse(dir.exists("../data/"),FALSE,dir.create("../data"))
ifelse(!file.exists("../data/Coursera-SwiftKey.zip"),download.file("https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip","../data/Coursera-SwiftKey.zip"),FALSE)
ifelse(dir.exists("../data/final/"),FALSE,unzip("../data/Coursera-SwiftKey.zip",exdir = file.path("../data/"),overwrite = FALSE))

if(!file.exists("data_final.RData")){
  data_path = list(US_Blog = "../data/final/en_US/en_US.blogs.txt",
                 US_Twitter = "../data/final/en_US/en_US.twitter.txt",
                 US_News = "../data/final/en_US/en_US.news.txt")
  data = map(data_path,readLines)
  data <- c(data$US_Blog,data$US_Twitter,data$US_News)
  save(data,file = "data_final.RData")
}else{
  if(!file.exists("sampled_data_clean.RData")){
    load("data_final.RData")
    }
}
```

The basic summary of the data is as follows:

## Data Preprocessing
Following pre-processing is done on the data:

* Data is sampled as data is too large. Only 3 % of the data is taken from each 
data file.
* Corpus is generated
* Following transformations are performed:

    1. NonASCII characters are removed.
    2. Profanity words are removed.
    3. StopWords are removed.
    4. URLs are removed.
    5. Punctuations are removed.
    6. Numbers are removed.
    7. Words are uncapitalized.
    8. Spaces are removed.
  
* Creating DocumentTermMatrices:
    1. 1-Gram
    2. 2-Gram
    3. 3-Gram
    4. 4-Gram

```{r clean_corpus,eval=FALSE,echo=FALSE,cache=TRUE,warning=FALSE}
# Clean Data
profanity <- readLines("https://www.cs.cmu.edu/~biglou/resources/bad-words.txt")
clean.text <- function(doc,profanity) {
  library(tm)
  doc <- iconv(doc, "latin1", "ASCII", sub="")
  # "won't" is a special case as it does not expand to "wo not"
  doc <- gsub("won't", "will not", doc)
  doc <- gsub("can't", "can not", doc)
  doc <- gsub("n't", " not", doc)
  doc <- gsub("'ll", " will", doc)
  doc <- gsub("'re", " are", doc)
  doc <- gsub("'ve", " have", doc)
  doc <- gsub("'m", " am", doc)
  doc <- gsub("'d", " would", doc)
  # 's could be 'is' or could be possessive: it has no expansion
  doc <- gsub("'s", "", doc)
  doc <- tolower(doc)
  
  doc <- removeWords(doc,profanity)
  #doc <- gsub(prof_reg1,'',doc)
  #doc <- gsub(prof_reg2,'',doc)
  
  doc <- gsub("http[[:alnum:][:punct:]]*","",doc)
  doc <- gsub("[[:punct:]]*","",doc)
  doc <- gsub("[[:digit:]]*","",doc)
  doc <- gsub("\\s+"," ",doc)
  
  return(doc)
}
```

```{r}
set.seed(1234)
create_sample <- function(x){
  sample(data,length(data)* 0.01)
}

forced <- FALSE

# fix (expand) contractions
if(!file.exists("sampled_data_clean.RData") || forced){
  sampled_data <- map(1:100,create_sample)
  names(sampled_data) <- 1:100  
  opb <- pboptions(type="txt")
  
  # cleaning the samples
  sampled_data_clean <- pblapply(sampled_data,clean.text,
                             profanity,cl=cl)
  
  # creating the data frame
  sampled_data_clean <- pblapply(sampled_data_clean,
                             function(x){
                              library(tidyverse)  
                              data_frame(sentences=x)
                              },
                             cl=cl)
  
  save(sampled_data_clean,file = "sampled_data_clean.RData")

} else{
    load("sampled_data_clean.RData")
}


```


```{r}
create_ngram <- function(ngram=1){
  
  create_ngram_sampled <- function(data,n=1){
    library(tidytext)
    library(tidyverse)
    data %>% unnest_tokens(dgram,sentences,token = "ngrams", n = ngram) %>%
                  count(dgram, sort = TRUE)
  }
  
  opb <- pboptions(type="txt")
  dgram_sampled <- pblapply(sampled_data_clean,create_ngram_sampled,ngram,cl=cl)
  
  dgram <- bind_rows(dgram_sampled) %>% 
              group_by(dgram) %>%
              summarise(count = mean(n)) %>%
              filter(count>1) %>% arrange(desc(count)) %>%
              separate(dgram,sapply(1:ngram,function(x){paste0('word',x)}),sep = ' ')
  
  dgram
  
  
} 

```

### Bigram
```{r}
if(!file.exists("bigram.RData") || forced){
  bigram <- create_ngram(2)
  
  save(bigram,file = "bigram.RData")
}else{
  load("bigram.RData")
}
```

### Trigram
```{r}
if(!file.exists("trigram.RData") || forced){
  trigram <- create_ngram(3)
  
  save(trigram,file = "trigram.RData")
}else{
  load("trigram.RData")
}
```

### Fourgram
```{r}
set.seed(1234)
sample_50 <- sample(1:100,50)
if(!file.exists("fourgram.RData") || forced){
  if(!file.exists("fourgram_1.RData") || forced){
     
    fourgram_1 <- create_ngram(4,sampled_data_clean[sample_50])
    
    save(fourgram_1,file = "fourgram_1.RData")
  }else{
    load("fourgram_1.RData")
  }
  if(!file.exists("fourgram_2.RData") || forced){
    fourgram_2 <- create_ngram(4,sampled_data_clean[-sample_50])
    
    save(fourgram_2,file = "fourgram_2.RData")
  }else{
    load("fourgram_2.RData")
  }
  fourgram <- bind_rows(fourgram_1,fourgram_2) %>% group_by(word1,word2,word3,word4) %>%
              summarise(count = mean(count)) %>%
              arrange(desc(count))
  save(fourgram,file='fourgram.RData')
}else{
  load('fourgram.RData')
}
```

### Prediction
```{r}
predict <- function(input,top= 5){
  clean <- clean.text(input,profanity)
  words <- strsplit(clean,split = ' ')
  words <- unlist(words)
  print("Bigram Prediction:")
  bigram_pred <- bigram %>% filter(word1==words[length(words)]) %>% top_n(top)
  print(bigram_pred%>%select(2))
  print("Trigram Prediction")
  
  trigram_pred <- trigram %>% 
                      filter(word1==words[length(words)-1],
                             word2==words[length(words)]) %>%   
                      top_n(top)
  print(trigram_pred%>%select(3))
  
  print('FourGram Prediction')
  
  fourgram_pred <- fourgram %>% 
                      filter(word1==words[length(words)-2],
                             word2==words[length(words)-1],
                             word3==words[length(words)]) %>%   
                      top_n(top)
  
  print(fourgram_pred%>%select(4))
}
```