---
title: "The Analysis on the effect of Severe Weather on Public Health and Economy"
author: "Naqeeb Asif"
date: "1 February 2018"
output: html_document
---

# 1. Synopsis
Severe weather events bring catastrophe with them. They can be drastic to human
lives and can cause great damage to the economy. So it is better to be prepared
for the severe weather events beforehand and take extra precautions for the events
which do the most damage.<br> 
The aim of this project is to explore U.S. National Oceanic and Atmospheric 
Administration’s (NOAA) storm database and find out which event causes the most 
damage to the human lives and economy.<br>
After analysing the database it was found out that the most damage to the human lives 
was done by tornadoes and the most damage done to the economy was by the floods.


# 2. Obtaining the data
## 2.1. Data
The data for the anlaysis is in the form of a comma-separated-value file which is 
compressed via the bzip2 algorithm to reduce its size. The data can be downloaded
from the course website: [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2)
<br>
The documentation of the data is also available and can be donwloaded from 
[National Weather Service Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf).
The answers to some frequently asked questions are also avaible and can be 
download from [National Climatic Data Center Storm Events FAQ](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2FNCDC%20Storm%20Events-FAQ%20Page.pdf)
<br>
The database contains the recorde from year 1950 to Novemeber 2011. Due to the 
lack of good records there are fewere events record in many earlier years of the 
database. The record with the more recents year should be considered more complete.

## 2.2. Loading the required libraries
In this section the libraries required for the whole data are loaded. 

```{r load_libraries,echo=TRUE,results='hide',warning=FALSE,message=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(lubridate)
library(ggstance)
library(grid)
library(gridExtra)
```

## 2.3. Loading the data
The data is first downloaded from the course website. It is then loaded into R.

```{r load_data,echo=TRUE,results='hide',warning=FALSE,message=FALSE,cache=TRUE}

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2",
              destfile = "./FStormData.csv.bz2")
weather_data <- data.table(read.csv("./FStormData.csv.bz2"))
```

# 3. Data Processing

This section illustrates the steps performed on the original data to get the required
data only.

## 3.1. Data for Analyzing Effects on Population Health

This section illustrates how data for analyzing effects on population health was
created. Three columns namely "EVTYPE","FATALITIES" and "INJURIES" were selected
from the original data. The subset was then grouped by "EVETYPE" and for each event
in "EVTYPE" total fatalities and total injuries were calculated and stored in 
health_data table.  

After creating health_data its two subset were created named fatalities_data and
injuries_data which contain top10 total fatalities and top10 total injuries 
respectively.

```{r data_health,echo="TRUE",results="hide",results='hide',message=FALSE}

health_data <- weather_data %>% select(EVTYPE,FATALITIES,INJURIES) %>% 
               group_by(EVTYPE) %>% summarise(total_fatalities=sum(FATALITIES),
                                              total_injuries=sum(INJURIES))

fatalities_data <- health_data %>% select(-total_injuries) %>%
                   arrange(desc(total_fatalities)) %>% top_n(10)

injuries_data <- health_data %>% select(-total_fatalities) %>%
                   arrange(desc(total_injuries)) %>% top_n(10)

```

## 3.2. Data for Analyzing effects on Economy

This section illustrates how data for analyzing effects on population health was
created. Five columns namely "EVTYPE","PROPDMG","PROPDMGEXP","CROPDMG" and "CROPDMGEXP"
were selected from the original data. Two variables were introduced namely property_damge
and crop_damage. There values were calculated PROPDMG,PROPDMGEXP and CROPDMG,CROPDMGEXP
respectively. The values in PROPDMGEXP and CROPDMG were converted to number using
exp_to_number function. Function exp_to_number converts properly according to the 
values in PROPDMGEXP CROPDMG. The subset of the resultant data is then created using
three columns "EVTYPE","property_damage" and "crop_damage" which is then group by
EVETYPE and for each event in EVTYPE total property damage, total crop damage and 
total damage are then calculated. The resultant dataset is then stored in the table
economy_data.

After creating economy_data its three subset were created named property_data,
crop_data and total_ec_data were created which contain top10 total property ,
top10 total crop and top 10 total damgages respectively. 

```{r data_economy,echo="TRUE",results="hide",message=FALSE}

exp_to_number = function(x){
  v <- rep(0,length(x))
  v[x %in% c("H","h") ] = 10^2
  v[x %in% c("K","k") ] = 10^3
  v[x %in% c("M","m") ] = 10^6
  v[x %in% c("b","B") ] = 10^9
  
  v
}

economy_data <- weather_data %>% select(EVTYPE,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP) %>%
                mutate(property_damage = PROPDMG*exp_to_number(PROPDMGEXP),
                       crop_damage = CROPDMG * exp_to_number(CROPDMGEXP)) %>%
                select(EVTYPE,property_damage,crop_damage) %>%
                group_by(EVTYPE) %>%summarise(total_property_dmg = sum(property_damage),
                                            total_crop_dmg=sum(crop_damage)) %>%
                mutate(total_damage=total_property_dmg+total_crop_dmg)

property_data <- economy_data %>% select(EVTYPE,total_property_dmg) %>%
                 arrange(desc(total_property_dmg)) %>%top_n(10) 

crop_data <- economy_data %>% select(EVTYPE,total_crop_dmg) %>%
                 arrange(desc(total_crop_dmg)) %>%top_n(10) 

total_ec_data <- economy_data %>% select(EVTYPE,total_damage) %>%
                 arrange(desc(total_damage)) %>%top_n(10) 

```

# 4. Results
In this section results were produced according to the questions asked.

## 4.1. Imapct on Health

In this section the graph for predicting the imapct of severe weather events on 
health was plotted.

```{r results_health}
pal <- colorRampPalette(c("green","red"))

g1 <- ggplot(data=fatalities_data,aes(x=reorder(EVTYPE,total_fatalities),y=total_fatalities))+
      geom_bar(stat = "identity",fill=pal(10)) +ylab("Total Fatalities") +xlab("Event Type")+
      theme(plot.title = element_text(hjust = 0.5))+coord_flip() + 
      labs(title="Impact of Severe Weather Events on Health")

g2 <- ggplot(data=injuries_data,aes(x=reorder(EVTYPE,total_injuries),y=total_injuries))+
      geom_bar(stat = "identity",fill=pal(10)) +ylab("Total Injuries") +xlab("Event Type")+
      theme(plot.title = element_text(hjust = 0.5))+coord_flip()

grid.arrange(g1,g2,ncol=1)
```


### 4.1.1 Analysis
From the plot it can be seen that tornadoes caused the total fatalities as well as 
total injuries in U.S.

## 4.2. Impact on Economy

In this section the graph for predicting the imapct of severe weather events on 
economy was plotted. Two graph were plotted in this section. One graph gives the 
damage done to property and crops and other graph gives the damage done to the 
overall economy.

### 4.2.1 Damage done to Property and Crops

```{r results_property_crops}
pal <- colorRampPalette(c("green","red"))

p1 <- ggplot(data=property_data,aes(x=reorder(EVTYPE,total_property_dmg),
                                    y=total_property_dmg))+geom_bar(stat = "identity",
                                                                    fill=pal(10)) +
      ylab("Total Property Damage in Dollars($)") +xlab("Event Type")+
      theme(plot.title = element_text(hjust = 0.5))+coord_flip() + 
      labs(title="Impact of Severe Weather Events on Property and Crops")

p2 <- ggplot(data=crop_data,aes(x=reorder(EVTYPE,total_crop_dmg),y=total_crop_dmg))+
      geom_bar(stat = "identity",fill=pal(10)) +ylab("Total Crop Damage in Dollars($)") +
      xlab("Event Type")+coord_flip()


grid.arrange(p1,p2,ncol=1)
```

### 4.2.2 Damage done to Overall Economy

```{r total_resut}

p3 <- ggplot(data=total_ec_data,aes(x=reorder(EVTYPE,total_damage),y=total_damage))+
      geom_bar(stat = "identity",fill=pal(10)) +ylab("Total Damage to Economy in Dollars($)") +
      xlab("Event Type")+theme(plot.title = element_text(hjust = 0.5))+coord_flip() + 
      labs(title="Impact of Severee Weather Events on Overall Economy")

p3
```

### 4.2.3 Analysis
From the plot it can be seen floods caused the most damage to the property and 
drought caused the most damage to the crops. Overall the floods caused the most 
damage to the economy.

#5. Conclusion
From the results it can be concluded that the toranado caused the most damage to
the injuries and people lives i.e. tornadoes impacted the human lives the most in United
States. Wheras the floods casued the most damage to the property and draughts caused the most
damage to the crops. Overall floods impacted the economy the most in United States.