---
title: "Week2"
author: "Naqeeb Asif"
date: "17 February 2018"
output: html_document
---

#Linear Regression for Prediction

##Example 
### Diamond data set from 'UsingR'
```{r,message=FALSE}
library(UsingR)
data(diamond)
library(ggplot2)
g <- ggplot(diamond,aes(x=carat,y=price))
g <- g +xlab("Mass (carats)")
g <- g + ylab("Price (SIN $)")
g <- g + geom_point(size=6,colour="black",alpha=0.2)
g <- g + geom_point(size=5,colour="blue",alpha=0.2)
g <- g + geom_smooth(method="lm",colour="black")
g

```

## Fitting the linear regression model

```{r}
fit <- lm(price~carat,data = diamond)
coef(fit)
```

## Getting a more interpretable intercept
```{r}
fit2 <- lm(price~I(carat-mean(carat)),data=diamond)
coef(fit2)
```

500$ is the expected price for the average sized diamond of the data.

## Predicting the price of the diamond
```{r}
newx <- c(0.16,0.27,0.34)
coef(fit)[1] + coef(fit)[2]*newx
predict(fit,newdata=data.frame(carat=newx))
```


# Residual Examples
## Residual Examples
```{r}

data(diamond)
fit <- lm(price~carat,data=diamond)
e <- resid(fit)
diamond$residuals <- e
```

## Residual plots
```{r}
library(ggplot2)
g <- ggplot(data=diamond,aes(y = residuals,x = carat))
g <- g + xlab("Mass (carats)") + ylab("Residuals (SIN$)") 
g <- g + geom_point(color="black",size=2,alpha=0.5)+geom_hline(yintercept = 0)
g +geom_segment(aes(xend=carat, yend=0), color="red")
```

## Non-linear data
```{r,message=FALSE}
x <- runif(100,-3,3)
y <- x +sin(x) + rnorm(100,sd=.2)
library(ggplot2)
g <- ggplot(data.frame(x =x,y=y),aes(x=x,y=y))
g <- g + geom_smooth(method = "lm",color="black")
g <- g + geom_point(size=7,color="black",alpha=0.5)
g <- g + geom_point(size=5,color="red",alpha=0.4)
g
```
### Residual Plot
```{r,message=FALSE}
library(ggplot2)
g <- ggplot(data.frame(x =x,y=resid(lm(y~x))),aes(x=x,y=y))
g <- g + geom_smooth(method = "lm",color="black")
g <- g + geom_point(size=7,color="black",alpha=0.5)
g <- g + geom_point(size=5,color="red",alpha=0.4)
g
```

### Daimond Data residual plots

```{r}
data(diamond)
e <- c(resid(lm(price~1,data=diamond)),
       resid(lm(price~carat,data=diamond)))
fit <- factor(c(rep("Itc",nrow(diamond)),
              rep("Itc,slope",nrow(diamond))))
g <- ggplot(data.frame(e=e,fit=fit),aes(y=e,x=fit,fill=fit))
g <- g + geom_dotplot(binaxis = "y",size=2,stackdir = "center")
g <- g +xlab("Fitting Approach") + ylab("Residual price")
g

```

## Residual Variance
```{r}
fit <- lm(price~carat,data=diamond)
summary(fit)$sigma
```

```{r}
n <- nrow(diamond)
sqrt(sum(resid(fit)^2)/(n-2))
```


# Ingerence in regression
## Example
```{r}
library(UsingR)
data("diamond")
y <- diamond$price
x <- diamond$carat
n <- length(y)
beta1 <- cor(x,y) * sd(y)/sd(x)
beta0 <- mean(y) - beta1 * mean(x)
e <- y - beta0 - beta1*x
sigma <- sqrt(sum(e^2))/n-2
ssx <- sum((x-mean(x))^2)
seBeta0 <- (1/n + mean(x)^2/ssx) ^ 0.5 * sigma
seBeta1 <- sigma/sqrt(ssx)
tBeta0 <- beta0/seBeta0; tBeta1 <- beta1/seBeta1
pBeta0 <- 2 *pt(abs(tBeta0),df=n-2,lower.tail = FALSE)
pBeta1 <- 2 *pt(abs(tBeta1),df=n-2,lower.tail = FALSE)
coefTable <- rbind(c(beta0,seBeta0,tBeta0,pBeta0),
                   c(beta1,seBeta1,tBeta1,pBeta1))
coefTable
colnames(coefTable) <- c("Estimate","Std.Error","t value","P(>|t|")
rownames(coefTable) <- c("(Intercept)","x")
coefTable

```

```{r}
fit <-lm(y~x)
summary(fit)$coefficients
```

## Prediction

```{r}
library(ggplot2)
newx = data.frame(x = seq(min(x), max(x), length = 100))
p1 = data.frame(predict(fit, newdata= newx,interval = ("confidence")))
p2 = data.frame(predict(fit, newdata = newx,interval = ("prediction")))
p1$interval = "confidence"
p2$interval = "prediction"
p1$x = newx$x
p2$x = newx$x
dat = rbind(p1, p2)
names(dat)[1] = "y"
g = ggplot(dat, aes(x = x, y = y))
g = g + geom_ribbon(aes(ymin = lwr, ymax = upr, fill = interval), alpha = 0.2)
g = g + geom_line()
g = g + geom_point(data = data.frame(x = x, y=y), aes(x = x, y = y), size = 4)
g
```

# Quiz
## Q1


```{r}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
summary(lm(y~x))
```

## Q2
```{r}
summary(lm(y~x))$sig
```

## Q3

```{r}
data("mtcars")
fit <- lm(mpg~wt,data=mtcars)
predict(fit,newdata = data.frame(wt=mean(mtcars$wt)),interval = ("confidence"))
```

## Q4
```{r}
help(mtcars)
```

##Q5

```{r}
data("mtcars")
fit <- lm(mpg~wt,data=mtcars)
predict(fit,newdata = data.frame(wt=3),interval = ("prediction"))

```

## Q6

```{r}
data("mtcars")
fit <- lm(mpg~wt,data=mtcars)
predict(fit,newdata = data.frame(wt=2),interval = ("confidence"))-
  predict(fit,newdata = data.frame(wt=0),interval = ("confidence"))

```

##Q9
```{r}
data("mtcars")
fit1 <- lm(mpg~wt,data=mtcars)
fit2 <-lm(mpg~1,data=mtcars)
sum(summary(fit1)$sigma ^ 2)/sum(summary(fit2)$sigma ^ 2)
```